A survey on NoSQL and NewSQL datastores

### 2014-12-09 paper_dec/paper_dec.tex
added Mongo & Impala to conclusion

VoltDB is a relational in-memory distributed database, aiming to couple the guarantees of classical SQL-stores with the scaling of NoSQL systems while performing at very high speeds\cite{stonebraker2013voltdb}.
\\VoltDB stores data in the traditional relational model, but replicated and partitioned (using consistent hashing) over several nodes\cite{grolinger2013data}. The data is queryable through a (growing) subset of SQL-92 \cite{voltdb2010voltdb}. Queries are preferably defined as stored procedures written in Java, in which the SQL statements are embedded. VoltDB supports primary and secondary indexing, and leaves the user the choice between hash- and tree-indexes\cite{voltdb_indexes}. VoltDB plans and optimizes queries in stored procedures optimized at compile time\cite{voltdb_query_plans}.\\
Relying entirely on DRAM makes it expensive to scale to petabyte scale datavolumes, but VoltDB can export data to other, more suited DMBSs such as columnar NoSQL stores.
\\In terms of concurrency control, VoltDB supports full ACID-transactions which execute simultaneously on all replicas. Main memory is divided into chunks and these are statically assigned to individual, single-threaded cores. A global controller serializes all multi-node transactions to a sequence of single-node transactions and inserts those into the transactions queues of the respective nodes. In this way, VoltDB obviates the need for locking and latching techniques.
\\\\
In short, VoltDB provides relational in memory storage for relatively large datasets, with very fast SQL-query capacities and ACID transactions. It is thus more suited for compute intensive applications that don't work on exorbitantly large data volumes but require very low latency.

@@

Cloudera Impala is a distributed SQL-engine running on top of the Hadoop stack, either on HDFS or HBase \cite{cloudera_impala}. It is intended specifically for analytical use, focussing on delivering real-time querying capacities and not on high-throughput on writes.
\\Data is accessible through a subset of SQL-92.
\\Impala's architecture is almost perfectly symmetrically distributed: all nodes execute the same \texttt{impalad}-process, which is responsible for the main database functionality. Queries can be sent to any node which will then function as coordinator for the specific query. There are however two processes that run on only one (not necessarily the same) node in the cluster, performing bookkeeping tasks and relaying metadata changes through the cluster\cite{impala_components}.
\\Contrary to the options discussed before, Impala doesn't partition rows by default. It does give the user the possibility to partition data should the data size call for it\cite{impala_partitioning}.
\\Impala has a highly efficient I/O-layer keeping disk- and CPU-utilization high at all times, resulting in considerably faster performance than other SQL-on-Hadoop solutions such as Apache Hive \cite{floratou2014sql}. However, this comes with the drawback that the working set of a query has to fit in the aggregate physical memory of the cluster it runs on. This puts some restrictions on the size of datasets processible with Impala, not fully using the scaling capacity of the underlying Hadoop layer.
\\Because of its analytical purposes, Impala doesn't have extensive concurrency control features, relying instead on the underlying storage layer. With HBase's excellent concurrency control features, this doesn't necessarily pose a problem.

@@

In recent years, progress in microbiology and bioinformatics has driven down the cost of DNA-sequencing. As DNA of more and more organisms is being sequenced, the question arises to make this process ever more efficient and to store the generated data in a scalable, performant and accessible way. This study has focused on NoSQL and NewSQL datastores to be used in both the sequencing pipeline as in the storage and analysis of the results. These solutions offer elastic scalability, flexible data modelling, good behavior in a distributed setting and resilience to failures.\\
Specifically, this paper has reviewed 6 different datastores, 2 of the most popular ones in three categories, and compared them on a selection of properties relevant to HPC applications. Columnar stores like Apache Cassandra and HBase are well suited to the task of storing extremely large datasets in a reliable and performant way, and could handle the storage of already sequenced genomes. VoltDB and CouchBase Server offer low latency on reads and writes, albeit (especially VoltDB) on smaller datasets, but would thus be very useful for storing rapidly changing intermediate data in the sequencing pipeline. %TODO quod Impala & MongoDB?

>>>>

VoltDB is a relational in-memory distributed database, aiming to couple the guarantees of classical SQL-stores with the scaling of NoSQL systems while performing at very high speeds\cite{stonebraker2013voltdb}.
\\VoltDB stores data in the traditional relational model, but replicated and partitioned (using consistent hashing) over several nodes\cite{grolinger2013data}. The data is queryable through a (growing) subset of SQL-92 \cite{voltdb2010voltdb}. Queries are preferably defined as stored procedures written in Java, in which the SQL statements are embedded. VoltDB supports primary and secondary indexing, and gives the user the choice between hash- and tree-indexes\cite{voltdb_indexes}. VoltDB plans and optimizes queries in stored procedures optimized at compile time\cite{voltdb_query_plans}.\\
Relying entirely on DRAM makes it expensive to scale to petabyte scale datavolumes, but VoltDB can export data to other, more suited DMBSs such as columnar NoSQL stores.
\\In terms of concurrency control, VoltDB supports full ACID-transactions which execute simultaneously on all replicas. Main memory is divided into chunks and these are statically assigned to individual, single-threaded cores. A global controller serializes all multi-node transactions to a sequence of single-node transactions and inserts those into the transactions queues of the respective nodes. In this way, VoltDB obviates the need for locking and latching techniques.
\\\\
In short, VoltDB provides relational in memory storage for relatively large datasets, with very fast SQL-query capacities and ACID transactions. It is thus more suited for compute intensive applications that don't work on exorbitantly large data volumes but require very low latency.

@@

Cloudera Impala is a distributed SQL-engine running on top of the Hadoop stack, either on HDFS or HBase \cite{cloudera_impala}. It is intended specifically for analytical use, focussing on delivering real-time querying capacities and not on high-throughput on writes.
\\Data is accessible through a subset of SQL-92.
\\Impala's architecture is almost perfectly symmetrically distributed: all nodes execute the same \texttt{impalad}-process, which is responsible for the main database functionality. Queries can be sent to any node which will then function as coordinator for the specific query. There are however two processes that run on only one (not necessarily the same) node in the cluster, performing bookkeeping tasks and relaying metadata changes through the cluster\cite{impala_components}.
\\Contrary to the options discussed before, Impala doesn't partition rows by default. It does give the user the possibility to partition data should the data size call for it\cite{impala_partitioning}.
\\Impala has a highly efficient I/O-layer keeping disk- and CPU-utilization high at all times, resulting in considerably faster performance than other SQL-on-Hadoop solutions such as Apache Hive \cite{floratou2014sql}. However, this comes with the drawback that the working set of a query has to fit in the aggregate physical memory of the cluster it runs on. This puts some restrictions on the size of datasets processable with Impala, not fully using the scaling capacity of the underlying Hadoop layer.
\\Because of its analytical purposes, Impala doesn't have extensive concurrency control features, relying instead on the underlying storage layer. With HBase's excellent concurrency control features, this doesn't necessarily pose a problem.

@@

In recent years, progress in microbiology and bioinformatics has driven down the cost of DNA-sequencing. As DNA of more and more organisms is being sequenced, the question arises to make this process ever more efficient and to store the generated data in a scalable, performant and accessible way. This study has focused on NoSQL and NewSQL datastores to be used in both the sequencing pipeline as in the storage and analysis of the results. These solutions offer elastic scalability, flexible data modelling, good behavior in a distributed setting and resilience to failures.\\
Specifically, this paper has reviewed 6 different datastores, 2 of the most popular ones in three categories, and compared them on a selection of properties relevant to HPC applications. Columnar stores like Apache Cassandra and HBase are well suited to the task of storing extremely large datasets in a reliable and performant way, and could handle the storage of already sequenced genomes. MongoDB sets itself apart with its extensive API, flexible data modeling, querying and indexing features. It doesn't have quite the write-handling capacities of the columnar stores, but is nevertheless a strong candidate for storing very large datasets such as the sequenced genomes. VoltDB and CouchBase Server offer low latency on reads and writes, albeit (especially VoltDB) on smaller datasets, but would thus be very useful for storing rapidly changing intermediate data in the sequencing pipeline. Lastly, Cloudera Impala lends itself well to situations where fast read, but not write queries on large datasets are required.
